---
title: "GAE0 -- Introduction to using GitHub Classroom"
author: "TH Martin"
date: "Spring 2021"
output:
  pagedown::html_paged:
    toc: true
    number_sections: false
    self_contained: true
    css: "theme.css"
  html_notebook:
bibliography: 
- GAE0.bib
- R-GAE0.bib
csl: ecology.csl
# uncomment this line to produce HTML and PDF in RStudio:
#knit: pagedown::chrome_print
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(papaja)
library(gtsummary)
library(tidyverse)

# Use the papaja package to create a bibliography of package citations:
papaja::r_refs(file = "R-GAE0.bib")

```

# Scenario 1

Relatively early in your textbook [@QK], you'll be introduced to an example data set that was generated by @lovettetal2000. They studied the chemistry of forested watersheds in the Catskill Mountains in New York State. We'll use this data file, ***lovett.csv*** as an introduction to using latex to write symbols and equations and to become more familiar with R, RStudio, and github.

@lovettetal2000 chose 39 sites (observations) on first and second order streams and measured the concentrations of ten chemical variables ($NO_3^-$, total organic $N$, total $N$, $NH_4^2$, dissolved organic $C$, $SO_4$ , $Cl^-$, $Ca^{2+}$, $Mg^{2+}$ , $H^+$), averaged over three years, and four watershed variables (maximum elevation, sample elevation, length of stream, watershed area). We will assume that the 39 sites represent a random sample of possible sites in the central Catskills and will focus on producing a table of summary statistics with only the stream name (*STREAM*), elevation (*ELEVATN*), $SO_4$ (*SO4*), and $Cl^-$ (CL) in our data file.

## Statistical Methods:

We will use the gtsummary package [@R-gtsummary] to produce a formatted table that reports the mean and standard deviation, median and interquartile range (IQR), and sample size for each of the chemical variables that were measured.

```{r dataImport}
df1 <- read_csv("lovett.csv")
```



# References:
